---
title: "Blank RMarkdown file."
author: "Jesse Leigh Patsolic"
output: 
  html_document:
    keep_md: true
    toc: true
    code_folding: show
    self_contained: true
---

<!--
### ### INITIAL COMMENTS HERE ###
###
### Jesse Leigh Patsolic 
### 2018 <jpatsolic@jhu.edu>
### S.D.G 
#
-->

```{r render-code, eval=FALSE, echo=FALSE}
set.seed(3171)
require(rmarkdown)
require(knitr)
knitr::opts_chunk$set(cache=FALSE, autodep=TRUE, eval = TRUE)

rmarkdown::render("GaborFilter_01.Rmd")
system('open GaborFilter_01.html')
```

<style type="text/css">
.table {
    width: 40%;
}
tr:hover {background-color:#f5f5f5;}
</style>

```{r setup}
require(rmarkdown)
require(devtools)
require(raster)
require(ggplot2)
require(MCMCpack)
require(rerf)
require(data.table)
require(gridExtra)
require(rgl)
data(mnist)
source("RerF-gaborFilter01/R/RandMat.R")
```

# Introduction 

The Gabor filter as defined by Goodfellow et al. is given by 
$$ w(x,y; \alpha, \beta_x, \beta_y, f, \phi, x_0, y_0, \tau) = \alpha
\exp{(- \beta_x x'^2 - \beta_y y'^2 \color{red}{+ \beta_{xy}x'y'})} \cos{(fx' + \phi)} $$

the term in red is required if the correlation is non-zero, however
this term is missing from the definition in Goodfellow et al.

For reference the bivariate Gaussian is given by 

$$ f(x,y; \mu_x, \mu_y, \sigma_x, \sigma_y, \rho) = \frac{1}{2\pi\sigma_x\sigma_y(1 - \rho^2)}\exp{\left\{-\frac{1}{2(1 - \rho^2)}\left[\frac{(x - \mu_x)^2}{\sigma_x^2} \color{blue}{- \frac{2\rho(x - \mu_x)(y - \mu_y)}{\sigma_x\sigma_y}}+ \frac{(y - \mu_y)^2}{\sigma_y^2} \right]\right\}} $$

**The conjugate prior for a uni-variate normal with known mean is an inverse gamma.**



| |  0|  1|  2|  3|
|:--:|--:|--:|--:|--:|
|$\Sigma$    | NA           |  invW|  invW|  invW|
|$\beta_x$   | $\Gamma(1,1)$|  $1/(2\sigma_x^2(1-\rho^2))$| ''|$1/(2\sigma_x^2)$|
|$\beta_y$   | $\Gamma(1,1)$|  $1/(2\sigma_y^2(1-\rho^2))$| ''|$1/(2\sigma_y^2)$|
|$\beta_{xy}$ | 0            |  $\rho/(\sigma_x\sigma_y(1-\rho^2))$|  0|  0|



### Gabor filter weight function

```{r GaborFunctionDef, echo = FALSE, results = 'asis'}
Gw
```

### gabSamp function

```{r gabSamp, echo = FALSE, results = 'asis'}
gabSamp
```


```{r param-test, eval = TRUE, include = FALSE}
set.seed(1234)
n <- 64
w <- 28
h <- 28
sw <- seq(-3.5, 3.5, length = w)
sh <- seq(3.5, -3.5, length = h)
xy <- expand.grid(sw,sh)

x <- xy[, 1]
y <- xy[, 2]

#params <- list()

alpha <- 1
betax <- rinvgamma(1, 1, 0.5)#sample(seq(0.01, 0.5,by=0.001), 1)#rinvgamma(1, 1,0.75)
betay <- rinvgamma(1, 1, 0.5)#sample(seq(0.01, 0.5,by=0.001), 1)#rinvgamma(1, 1, 0.75) 
f <- runif(1, min = iw / 8, max = 8 * iw)
phi <- 0
x0 <- 0#runif(1, -3,3)
y0 <- 0#runif(1, -3,3) 
tau <- 0

#G1 <- do.call(w, params)

G1 <- Gw(x, y, alpha, betax, betay, f, phi, x0, y0, tau)
d1 <- cbind(xy,G1)

#ggplot(data = d1, aes(x = Var1, y = Var2, fill = G1)) + geom_raster() + scale_fill_distiller(palette = "Spectral")
pq <- ggplot(data = d1, aes(x = Var1, y = Var2, fill = G1)) + 
  geom_raster() + 
  scale_fill_gradient2(low = "blue", mid = 'white', high = 'red')
plot(pq)
```


# Sampling Gabor filters


Given a random patch with height `h` and width `w` how do we sample a corresponding Gabor filter?

```{r check-looks-like-gabor}
set.seed(32)
# plots list
p0 <- list()
im <- matrix(mnist$Xtrain[1,], 28, 28)
## Get a random subsample, 100 each of 3's and 5's
#set.seed(320)
#threes <- sample(which(mnist$Ytrain %in% 3), 100)
#fives  <- sample(which(mnist$Ytrain %in% 5), 100)
#numsub <- c(threes, fives)
#
#Ytrain <- mnist$Ytrain[numsub]
#Xtrain <- mnist$Xtrain[numsub,]
#Ytest <- mnist$Ytest[mnist$Ytest %in% c(3,5)]
#Xtest <- mnist$Xtest[mnist$Ytest %in% c(3,5),]
#X <- Xtrain[1, ]
#
#for(i in sample(200)){
#  X <- Xtrain[i,]
#  plot(as.raster(matrix(X, 28, 28)))
#  Sys.sleep(0.5)
#  plot(raster::raster(matrix(X * G1, 28, 28)), col = gray.colors(255))
#  Sys.sleep(0.5)
#}

w <- 28

for(i in 1:25) {
  s  <- seq(-4, 4, length = w)

  xy <- expand.grid(s,s)
  
  param0 <- gabSamp(iw = 28)

  param0$x <- xy[, 1]
  param0$y <- xy[, 2]

  G0 <- do.call(Gw, param0)
  d0 <- data.table(cbind(xy,G0, GX = G0 * X))
  
  #ggplot(data = d1, aes(x = Var1, y = Var2, fill = G1)) + geom_raster() + scale_fill_distiller(palette = "Spectral")
  p0[[i]] <- ggplot(data = d0, aes(x = Var1, y = Var2, fill = G0)) + geom_raster() + scale_fill_gradient2(low = "blue", mid = 'white', high = 'red') + guides(fill = FALSE)

}
```


```{r, fig.height = 8, fig.width = 8}
#pdf('tmp.pdf', height = 10, width = 10)
print(do.call(grid.arrange, p0))
#dev.off()
```


### R-RerF RandMatImageGabor

```{r RandMatGabor-Def, echo = FALSE, results='asis'}
RandMatGabor
```

```{r mnist-sub}
## Get a random subsample, 100 each of 3's and 5's
set.seed(320)
zeros  <- sample(which(mnist$Ytrain %in% 0), 100)
ones   <- sample(which(mnist$Ytrain %in% 1), 100)
twos   <- sample(which(mnist$Ytrain %in% 2), 100)
threes <- sample(which(mnist$Ytrain %in% 3), 100)
eights <- sample(which(mnist$Ytrain %in% 8), 100)

numsub <- c(zeros, ones, eights)

Ytrain <- mnist$Ytrain[numsub]
Xtrain <- mnist$Xtrain[numsub,]

Ytest <- mnist$Ytest[mnist$Ytest %in% c(0, 1, 8)]
Xtest <- mnist$Xtest[mnist$Ytest %in% c(0, 1, 8),]

#s1 <- sample(nrow(Xtest))
#Ytrain <- mnist$Ytrain
#Xtrain <- mnist$Xtrain
#
#Ytest <- mnist$Ytest
#Xtest <- mnist$Xtest
```

## Run RerF with different parameter settings on a subset of MNIST
```{r randmatTest-setup}
# p is number of dimensions, d is the number of random features to evaluate, iw is image width, ih is image height, patch.min is min width of square patch to sample pixels from, and patch.max is the max width of square patch
p <- ncol(Xtrain)
d <- ceiling(sqrt(p))
iw <- sqrt(p)
ih <- iw
patch.min <- 2L
patch.max <- 10L

ntrees <- 500L
ncores <- 7L
seed1 <- 1L
```

### Parameters

```{r, results = 'asis'}
kable(as.data.frame((params <- list(p = p, d = d, iw = iw, ih = ih, pwMin = patch.min, pwMax = patch.max, sparsity = 1))))
```


### Image-Patch 

```{r RerF-image-patch}
system.time({
forestIP <- RerF(Xtrain, Ytrain,  FUN = RandMatImagePatch, 
               paramList = params,
               trees = ntrees, 
               num.cores = ncores, seed = seed1)
})

predIP <- Predict(Xtest, forestIP, num.cores = ncores)
(errorIP <- mean(predIP != Ytest))
```


### RerF Continuous

NOTE: that sparsity has been set to 1.

```{r RerF-Continuous}
#params <- list(p = p, d = d, sparsity = 1)
system.time({
forestC <- RerF(Xtrain, Ytrain,  FUN = RandMatContinuous, 
               paramList = params,
               trees = ntrees,
               num.cores = ncores, seed = seed1)
})

predC <- Predict(Xtest, forestC, num.cores = ncores)
(errorC <- mean(predC != Ytest))
```

### RerF Gabor

```{r RerF-Gabor}
#params <- list(p = p, d = d, iw = 28, ih = 28)
system.time({
forestG0 <- RerF(Xtrain, Ytrain,  FUN = RandMatImageGabor,
               paramList = params,
               trees = ntrees,
               num.cores = ncores, seed = seed1)
})

predG0 <- Predict(Xtest, forestG0, num.cores = ncores)
(errorG0 <- mean(predG0 != Ytest))
```

# results

```{r, echo = FALSE, results = 'asis'}
oe <- data.frame(
  cbind(
        Gabor0 = errorG0, 
        ImagePatch = errorIP, Continuous = errorC))
kable(oe[, order(oe[1, ])])
```






<!--
#   Time:
##  Working status:
### Comments:
####Soli Deo Gloria
--> 

