---
title: "Plot Gabor filters with varying parameters."
author: "Jesse Leigh Patsolic"
output: 
  html_document:
    keep_md: true
    toc: true
    code_folding: show
    self_contained: true
---

<!--
### ### INITIAL COMMENTS HERE ###
###
### Jesse Leigh Patsolic 
### 2018 <jpatsolic@jhu.edu>
### S.D.G 
#
-->

```{r render-code, eval=FALSE, echo=FALSE}
set.seed(3171)
require(rmarkdown)
require(knitr)
knitr::opts_chunk$set(cache=FALSE, autodep=TRUE, eval = TRUE)

rmarkdown::render("PlotGabor.Rmd")
system('open PlotGabor.html')
```

<style type="text/css">
.table {
    width: 40%;
}
tr:hover {background-color:#f5f5f5;}
</style>

```{r setup}
require(rmarkdown)
require(devtools)
require(raster)
require(ggplot2)
require(MCMCpack)
require(rerf)
require(data.table)
require(gridExtra)
require(rgl)
data(mnist)
source('RandMatGabor.R')
```

# Introduction 

The Gabor filter as defined by Goodfellow et al. is given by 
$$ w(x,y; \alpha, \beta_x, \beta_y, f, \phi, x_0, y_0, \tau) = \alpha
\exp{(- \beta_x x'^2 - \beta_y y'^2 \color{red}{+ \beta_{xy}x'y'})} \cos{(fx' + \phi)} $$

the term in red is required if the correlation is non-zero, however
this term is missing from the definition in Goodfellow et al.

For reference the bivariate Gaussian is given by 

$$ f(x,y; \mu_x, \mu_y, \sigma_x, \sigma_y, \rho) = \frac{1}{2\pi\sigma_x\sigma_y(1 - \rho^2)}\exp{\left\{-\frac{1}{2(1 - \rho^2)}\left[\frac{(x - \mu_x)^2}{\sigma_x^2} \color{blue}{- \frac{2\rho(x - \mu_x)(y - \mu_y)}{\sigma_x\sigma_y}}+ \frac{(y - \mu_y)^2}{\sigma_y^2} \right]\right\}} $$

**The conjugate prior for a uni-variate normal with known mean is an inverse gamma.**



| |  0|  1|  2|  3|
|:--:|--:|--:|--:|--:|
|$\Sigma$    | NA           |  invW|  invW|  invW|
|$\beta_x$   | $\Gamma(1,1)$|  $1/(2\sigma_x^2(1-\rho^2))$| $1/(2\sigma_x^2(1-\rho^2))$|$1/(2\sigma_x^2)$|
|$\beta_y$   | $\Gamma(1,1)$|  $1/(2\sigma_y^2(1-\rho^2))$| $1/(2\sigma_y^2(1-\rho^2))$|$1/(2\sigma_y^2)$|
|$\beta_{xy}$ | 0            |  $\rho/(\sigma_x\sigma_y(1-\rho^2))$|  0|  0|


## Plots varying $\beta_i$

```{r}
iw <- 28
ih <- 28
betax <- sort(rinvgamma(4, 1, 0.5))
betay <- sort(rinvgamma(4, 1, 0.5))

p <- expand.grid(alpha = 1, 
            betax = betax, 
            betay = betay, 
            f = 150, 
            phi = 0, x0 = 0, y0 = 0, tau = 0)

xg <- seq(-4,4, length = iw)
xy <- expand.grid(xg, xg)

pq <- list()

for(i in 1:nrow(p)){
  params <- c(p[i, ])
  params$x <- xy[, 1]
  params$y <- xy[, 2]

  #print(as.matrix(params[-c(9,10)]))
  G <- do.call(Gw, params)
  d1 <- data.table(cbind(xy, G))

  pq[[i]] <- ggplot(data = d1, aes(x = Var1, y = Var2, fill = G)) + 
    geom_raster() + 
    scale_fill_gradient2(low = "blue", mid = 'white', high = 'red') + 
    guides(fill = FALSE) + theme_void()
}

do.call(grid.arrange, pq)
```

## Plots varying $f$ and $\phi$ with $beta_i = 0.35$


```{r}
iw <- 28
ih <- 28
f <- runif(16, min = floor(iw/2), max = 10 * iw)#iw * seq(0.25, 10, length = 4)
phi <- 0#seq(0, 2 * pi, length = 4)

p <- expand.grid(alpha = 1, 
            betax = 0.35, 
            betay = 0.35, 
            f = f, 
            phi = phi, x0 = 0, y0 = 0, tau = 0)

xg <- seq(-4,4, length = iw)
xy <- expand.grid(xg, xg)

pq <- list()

for(i in 1:nrow(p)){
  params <- c(p[i, ])
  params$x <- xy[, 1]
  params$y <- xy[, 2]

  #print(as.matrix(params[-c(9,10)]))
  G <- do.call(Gw, params)
  d1 <- data.table(cbind(xy, G))

  pq[[i]] <- ggplot(data = d1, aes(x = Var1, y = Var2, fill = G)) + 
    geom_raster() + 
    scale_fill_gradient2(low = "blue", mid = 'white', high = 'red') + 
    guides(fill = FALSE) + theme_void()
}

do.call(grid.arrange, pq)
```

## Plots varying $x_0$, $y_0$, and $\tau$

```{r, fig.width = 10, fig.height = 10}
iw <- 28
ih <- 28

x0 <- seq(-2,2, length = 4)
y0 <- seq(2,-2, length = 4)
tau <- seq(-pi, pi, length = 4)

p <- expand.grid(alpha = 1, 
            betax = 0.35, 
            betay = 0.35, 
            f = 5 * iw, 
            phi = 0, x0 = x0, y0 = x0, tau = tau)

xg <- seq(-4,4, length = iw)
xy <- expand.grid(xg, xg)

pq <- list()

for(i in 1:nrow(p)){
  params <- c(p[i, ])
  params$x <- xy[, 1]
  params$y <- xy[, 2]

  #print(as.matrix(params[-c(9,10)]))
  G <- do.call(Gw, params)
  d1 <- data.table(cbind(xy, G))

  pq[[i]] <- ggplot(data = d1, aes(x = Var1, y = Var2, fill = G)) + 
    geom_raster() + 
    scale_fill_gradient2(low = "blue", mid = 'white', high = 'red') + 
    guides(fill = FALSE) + theme_minimal() + 
    theme(axis.title=element_blank(),
        axis.text=element_blank(),
        axis.ticks=element_blank())
}

do.call(grid.arrange, pq)
```



## Function code

### Gabor filter weight function

```{r GaborFunctionDef, echo = FALSE, results = 'markup'}
Gw
```

### gabSamp function

```{r gabSamp, echo = FALSE, results = 'markup'}
gabSamp
```


### RandMatGabor 
```{r, echo = FALSE, results = 'markup'}
RandMatGabor
```












<!--
#   Time:
##  Working status:
### Comments:
####Soli Deo Gloria
--> 

